---
title: "Retirement Income Toolkit - Economic Scenario - Pricing Kernels"
author: "Queenie Huang"
output:
 bookdown::html_document2:
    toc: yes
    toc_depth: '4'
    number_sections: yes
    fig_caption: yes
    code_folding: hide
urlcolor: blue
header-includes: |    
 \usepackage{amsmath,amsthm,amssymb,amsfonts, mathtools, empheq, bbm}
 \newcommand{\vect}[1]{\boldsymbol{#1}}
---

<style type="text/css">

body{ /* Normal  */
      font-size: 15px;
  }
td {  /* Table  */
  font-size: 15px;
}
h1.title {
  font-size: 30px;
}
h1 { /* Header 1 */
  font-size: 26px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 15px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 15px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
options(scipen = 1, digits = 6) # use 2 d.p. in output
```

```{r}
library(tidyverse)
library(readxl)
library(zoo)
library(janitor)
library(doParallel)
```


# Import Data 

```{r}
start_pc_date = as.Date("1993-04-01")
end_date = as.Date("2025-12-31")
max_maturity = 40
n_index = c(1,4,8,20,40)
maturities = c("Date", paste(1:max_maturity, "qtrs", sep = "_"))
var_names = c("zcp3m_yield", "zcp10y_spread", "home_index_growth", "GDP_growth", "CPI_growth", "ASX200_growth", "AUD_growth")

#######################
# import interest rates
zcp = read.csv("data/zcb.csv")[,c(1,n_index+1)]
names(zcp) = c("Date", "ZCP3m","ZCP1y","ZCP2y","ZCP5y","ZCP10y")
zcp$Date = as.Date(zcp$Date, format = "%d/%m/%Y")


# match the frequency 
zcp = zcp %>%
    filter(Date <= end_date & Date >= start_pc_date) %>%
    mutate(Year = year(Date), Quarter = quarter(Date)) %>%
    group_by(Year, Quarter) %>%
    filter(Date == max(Date)) %>%
    ungroup() %>%
    mutate(Qtr = as.yearqtr(Date)) %>%
    dplyr::select(Qtr, ZCP3m, ZCP1y, ZCP2y, ZCP5y, ZCP10y)

# add row names & change class()
time_names_past = zcp$Qtr
zcp_sdf = as.matrix(zcp[-1,-1]) / 100
zcp = as.matrix(zcp[-1,-1]) / 100
row.names(zcp) = as.character(time_names_past[-1])
row.names(zcp_sdf) = as.character(time_names_past[-1])

###################
# import zt vectors 
all_data = all_data_stationary
all_data = as.matrix(t(all_data))
row.names(all_data) = var_names
colnames(all_data) = as.character(time_names_past[-1])
all_data[1,] = all_data[1,]
all_data[2,] = all_data[2,]

##############
# VAR Dynamics
# intercept 
intercept = matrix(c_vec, nrow = 7, ncol = 1)
# coefficient matrix in VAR(1)
coef = data.frame(A1,
                  row.names = var_names)
colnames(coef) = var_names
coef = as.matrix(t(coef))

# residual covariance matrix 
covres = data.frame(Sigma,
                    row.names = var_names)
colnames(covres) = var_names
covres = as.matrix(t(covres))
simga_half = chol(covres)
e1 = as.matrix(c(1, rep(0,6)))
```

# Model - estimate lambda_0, lambda_1 (historical data)

```{r}
######################
# calculate An, Bn, Cn

get_ABC = function (lambda_0, lambda_1) {
  lambda_0 = as.matrix(lambda_0)
  lambda_1 = as.matrix(lambda_1)
  
  # initialise dataframes to store An, Bn, Cn
  An = matrix(NA, nrow = 1, ncol = max_maturity)
  Bn = matrix(NA, nrow = length(intercept), ncol = max_maturity)
  Cn = matrix(NA, nrow = length(intercept), ncol = max_maturity)
  
  # initial values A1, B1, C1
  An[1] = 0
  Bn[,1] = -e1
  Cn[,1] = rep(0,length(intercept))
  
  # terminal values A40, B40, C40
  # An[max_maturity] = 0
  # Bn[,max_maturity] = -max_maturity * e1 - max_maturity * c(0,1,rep(0,length(intercept)-2))
  # Cn[,max_maturity] = rep(0,length(intercept))
  # 
  # calculate recursively
  for (n in 1:(max_maturity-1)) {
    An[n+1] = An[n] + t(Bn[,n]) %*% (intercept - simga_half %*% lambda_0) + 1/2 * t(Bn[,n]) %*% covres %*% Bn[,n]
    Bn[,n+1] = -e1 + t(coef - simga_half %*% lambda_1) %*% Bn[,n] + Cn[,n]
    Cn[,n+1] = 0 %*% Bn[,n]
  }
  
  # output 
  output = list(An, Bn, Cn)
  names(output) = c("An", "Bn", "Cn")
  output = lapply(output, function(x){x = as.matrix(x)})
  output = lapply(output, function(x){colnames(x) = maturities[-1]; return (x)})
  return (output)
}

#############################################
# calculate sum (r(n)t hat - r(n)t) ^ 2 stats

# only for r(1), r(4), r(8), r(20), r(40)
get_rnt = function (lambda_0, lambda_1) {
  # An, Bn, Cn
  ABCn = get_ABC(lambda_0, lambda_1)
  An = ABCn$An
  Bn = ABCn$Bn
  Cn = ABCn$Cn

  # zt: all_data

  # initialise dataframe to store historical r_t^(n)
  rnt = matrix(NA, nrow = length(time_names_past)-1, ncol = max_maturity)
  row.names(rnt) = as.character(time_names_past[-1])
  colnames(rnt) = maturities[-1]

  # r(n)t hat
  for (t in 2:length(time_names_past)-1) {
    for (n in n_index) {
      rnt[t,n] = -An[n] / n - t(Bn[,n]) %*% as.matrix(all_data[,t]) / n
    }
  }

  return (rnt[,n_index])
}

# r(1), r(2), ..., r(40)
# get_rnt = function (lambda_0, lambda_1) {
#   # An, Bn, Cn
#   ABCn = get_ABC(lambda_0, lambda_1)
#   An = ABCn$An
#   Bn = ABCn$Bn
#   Cn = ABCn$Cn
#   
#   # zt: all_data
#   
#   # initialise dataframe to store historical r_t^(n)
#   rnt = matrix(NA, nrow = length(time_names_past), ncol = max_maturity)
#   row.names(rnt) = as.character(time_names_past)
#   colnames(rnt) = maturities[-1]
#   
#   # r(n)t hat  
#   for (t in 2:length(time_names_past)) {
#     for (n in 1:max_maturity) {
#       rnt[t,n] = -An[n] / n - t(Bn[,n]) %*% all_data[,t] / n - t(Cn[,n]) %*% all_data[,t-1] / n
#     }
#   }
#   
#   return (rnt[-1,])
# }

get_sse = function (lambda_list) {
  # reorganise parameters
  lambda_0 = matrix(lambda_list[1:7], nrow = 7)
  lambda_1 = matrix(lambda_list[8:56], nrow = 7)
  
  rnt_hat = get_rnt(lambda_0, lambda_1)
  sum((rnt_hat - zcp) ^ 2)
}

#################################
# minimise for lambda_0, lambda_1
lambda = rep(0, 8*7)

optim_ctrl = list(maxit = 500)
cl <- makeCluster(detectCores())
registerDoParallel(cl)
lambdas = optim(lambda, get_sse, control = optim_ctrl)
stopCluster(cl)
lambdas

lambda_0 = matrix(lambdas$par[1:7], nrow = 7)
lambda_1 = matrix(lambdas$par[8:56], nrow = 7)
```

# Model - historical s_t

```{r}
lambda_t = sapply(1:130, function (x) {lambda_0 + lambda_1 %*% all_data[,x]}, simplify = T)
st_past = sapply(1:130, function (x) {exp(-all_data[1,x] - 1/2 * sum(lambda_t[,x]^2))})
plot(time_names_past[-1], st_past * 100, "l", main = "Historical SDFs", xlab = "", ylab = "")
```


# Testing: get_var_simulations

```{r}
num_paths = 10
num_years = 20

# source("../SDF_Simulation.R")
# source("../VAR.R")

cl <- makeCluster(detectCores())
registerDoParallel(cl)
set.seed(1)
#sdf = get_var_simulations(num_paths = num_paths, num_years = num_years, return_sdf = T)
sdf = esg_var_simulator(num_paths = num_paths, num_years = num_years, return_sdf = T)
stopCluster(cl)

par(mfrow = c(2,2))

#######
# zcp3m
plot_date = c(as.Date(time_names_past)[-1], row.names(sdf$zcp3m_yield)[-1])
plot_data = c(all_data[1,], sdf$zcp3m_yield[-1,1])
plot(plot_date, plot_data, "l", main = "zcp3m", col = "grey")
for (i in 2:num_paths) {
  plot_data = c(all_data[1,], sdf$zcp3m_yield[-1,i])
  lines(plot_date, plot_data, col = "grey")
}

########
# zcp10y
plot_data = c(all_data[2,], sdf$zcp10y_spread[-1,1])
plot(plot_date, plot_data, "l", main = "zcp10y", col = "grey")
for (i in 2:num_paths) {
  plot_data = c(all_data[2,], sdf$zcp10y_spread[-1,i])
  lines(plot_date, plot_data, col = "grey")
}

#####
# sdf 
plot_data = c(st_past, sdf$discount_factors[-1,i])
plot(plot_date, plot_data, "l", main = "SDFs", xlab = "", ylab = "", col = "grey")
for (i in 2:num_paths) {
  plot_data = c(st_past, sdf$discount_factors[-1,i])
  lines(plot_date, plot_data, col = "grey")
}
```
