---
title: "Retirement Income Toolkit - Economic Scenario - VAR"
author: "Queenie Huang"
output:
 bookdown::html_document2:
    toc: yes
    toc_depth: '4'
    number_sections: yes
    fig_caption: yes
    code_folding: hide
urlcolor: blue
header-includes: |    
 \usepackage{amsmath,amsthm,amssymb,amsfonts, mathtools, empheq, bbm}
 \newcommand{\vect}[1]{\boldsymbol{#1}}
---

```{=html}
<style type="text/css">

body{ /* Normal  */
      font-size: 15px;
  }
td {  /* Table  */
  font-size: 15px;
}
h1.title {
  font-size: 30px;
}
h1 { /* Header 1 */
  font-size: 26px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 15px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 15px;
}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
options(scipen = 1, digits = 6) # use 2 d.p. in output
```

```{r}
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(readxl)
library(janitor)
library(lubridate)
library(vars)
library(tseries)
library(corrplot)
library(kableExtra)
library(reshape2)
library(latex2exp)
library(aTSA)
library(caret)
library(Metrics)
```

A VAR(p) process can be represented by \begin{align*}
    \vect Y_t = \vect \alpha + A_1 \vect Y_{t-1} + \ldots + A_p \vect Y_{t-p} + \vect \epsilon, 
\end{align*} where $n$ is the number of economic/financial variables, $p$ is the lag, $\vect \epsilon$ is a white noise process, $A_1, \ldots, A_p$ are fixed coefficient matrices.

# Economic/financial variables

```{r, message=FALSE, warning=FALSE}
# variables = read_excel("data/Variables.xlsx")

all_var_names = c("zcp3m", "zcp10y", "home_index", "GDP", "CPI", "ASX200", "forex") # those to be kept 
all_data_original = as.data.frame(matrix(NA, nrow = 131, ncol = 1+length(all_var_names)))
colnames(all_data_original) = c("Date", all_var_names)


# kbl(variables) %>% kable_styling(latex_options = c("striped", "scale_down"))
```

# Assess data quality & cleaning

## Data quality

- **Variable selection/transformation.**

  - `10-year zero-coupon spread` = `10-year zero-coupon yield` - `3-month zero-coupon yield`.

  - Negative `3-month zero-coupon yield`.

  - `Hedonic home value index` and `Hedonic rental yields` are medians of Sydney (SA4 regions) houses (excluding units).

- **Time window.** Q1 1993 - Q4 2025 (131 quarters).

- **Time frequency.** Converted to quarterly frequency by averaging.

- **Units.** Quarterly indexes are converted to quarterly rates.

  - `Hedonic home value growth rate`$_t$ = log(`Hedonic home value rate`$_t$) - log(`Hedonic home value rate`$_{t-1}$).

  - `GDP  growth rate`$_t$ = log(`GDP`$_t$) - log(`GDP`$_{t-1}$).

  - `S&P/ASX200 growth rate`$_t$ = log(`S&P/ASX200 price`$_t$) - log(`S&P/ASX200 price`$_{t-1}$).

  - `AUD trade index growth rate`$_t$ = log(`AUD index`$_t$) - log(`AUD index`$_{t-1}$).

```{r, message=FALSE, warning=FALSE}
start_pc_date = as.Date("1993-04-01")
start_index_date = as.Date("1993-1-01") # indexes are converted to percentages, one period short
end_date = as.Date("2025-12-31")

################################################
# ZERO-COUPON INTEREST RATES – ANALYTICAL SERIES
# import - 3 months 
# zcp_old = read_excel("data/ZCP_till2008.xls")[-c(1:10), c(1,3)]
# zcp_new = read_excel("data/ZCP_from2008.xls")[-c(1:10), c(1,3)]
# names(zcp_old) = c("Date", "ZCP3month")
# names(zcp_new) = c("Date", "ZCP3month")
# zcp3m = rbind(zcp_old, zcp_new)

zcp3m = read.csv("data/zcb.csv")[,c(1,2)]
names(zcp3m) = c("Date", "ZCP3month")

# import - 10 years 
# zcp_old = read_excel("data/ZCP_till2008.xls")[-c(1:10), c(1,42)]
# zcp_new = read_excel("data/ZCP_from2008.xls")[-c(1:10), c(1,42)]
# names(zcp_old) = c("Date", "ZCP10year")
# names(zcp_new) = c("Date", "ZCP10year")
# zcp10y = rbind(zcp_old, zcp_new)
# rm(zcp_new, zcp_old)

zcp10y = read.csv("data/zcb.csv")[, c(1,41)]
names(zcp10y) = c("Date", "ZCP10year")

# cleaning
# zcp3m = zcp3m %>% mutate_if(is.character, as.numeric)
# zcp10y = zcp10y %>% mutate_if(is.character, as.numeric)
# zcp3m$Date = as.Date(zcp3m$Date, origin = "1899-12-30")
# zcp10y$Date = as.Date(zcp10y$Date, origin = "1899-12-30")

zcp3m$Date = as.Date(zcp3m$Date, format = "%d/%m/%Y")
zcp10y$Date = as.Date(zcp10y$Date, format = "%d/%m/%Y")

# convert 10 year CZP to spread 
zcp10y$ZCP10y_spread = zcp10y$ZCP10year - zcp3m$ZCP3month

# match the time window 
zcp3m = zcp3m %>% filter(Date <= end_date & Date >= start_pc_date)
zcp10y = zcp10y %>% filter(Date <= end_date & Date >= start_pc_date)

# match the frequency 
zcp3m = zcp3m %>%
    mutate(Year = year(Date), Quarter = quarter(Date)) %>%
    group_by(Year, Quarter) %>%
    filter(Date == max(Date)) %>%
    ungroup() %>%
    mutate(Qtr = as.yearqtr(Date)) %>%
    dplyr::select(Qtr, ZCP3month) 

zcp10y = zcp10y %>%
    mutate(Year = year(Date), Quarter = quarter(Date)) %>%
    group_by(Year, Quarter) %>%
    filter(Date == max(Date)) %>%
    ungroup() %>%
    mutate(Qtr = as.yearqtr(Date)) %>%
    dplyr::select(Qtr, ZCP10y_spread, ZCP10year) 

all_data_original$Date = zcp10y$Qtr
all_data_original$zcp3m = zcp3m$ZCP3month
all_data_original$zcp10y = zcp10y$ZCP10year

# match the unit 

# declare as a ts object 
zcp3m = ts(zcp3m$ZCP3month)
#zcp3m = ts(log(zcp3m$ZCP3month_qtr_rate))
zcp10y_spread = ts(zcp10y$ZCP10y_spread)

###############
# Mortgage rate
lending = read_excel("data/Mortgage-RBA.xlsx", sheet = "Data")[-c(1:10), c(1,4)]
names(lending) = c("Date", "mortgage_rate")
lending = lending %>% mutate_if(is.character, as.numeric)
lending$Date = as.Date(lending$Date, origin = "1899-12-30")

# match the time window
lending = lending %>% filter(Date <= end_date & Date >= start_pc_date)

# match the frequency
lending = lending %>%
    mutate(Year = year(Date), Quarter = quarter(Date)) %>%
    group_by(Year, Quarter) %>%
    filter(Date == max(Date)) %>%
    ungroup() %>%
    mutate(Qtr = as.yearqtr(Date)) %>%
    dplyr::select(Qtr, mortgage_rate) 

# declare as a ts object
lending = ts(lending$mortgage_rate)

all_data_original$lending = lending

######################################
# House price dynamics
# import & cleaning 
properties = read.csv("data/home_price.csv")
properties = clean_names(properties)
properties = properties %>% mutate(Date = as.Date(date, format = "%d/%m/%Y")) %>% filter(Date <= end_date & Date >= start_index_date) %>% mutate(Qtr = as.yearqtr(Date))
properties = properties[nrow(properties):1, ]
all_data_original$home_index = properties$value_at_date[-1]


# match the unit 
home_value_houses_nsw = properties %>% mutate(home_value_qtr_rate = log(value_at_date) - log(dplyr::lag(value_at_date, n = 1))) 
home_value_houses_nsw = home_value_houses_nsw[-1,]

# declares variables as time_series objects 
home_value_houses_nsw = ts(home_value_houses_nsw$home_value_qtr_rate)

####################################################################
# Gross domestic product: Chain volume measures; seasonally adjusted
# import 
gdp_old = read_excel("data/GDP-ABS.xlsx", sheet = "Data1")[-c(1:9), c(1,54)]
gdp = read_excel("data/GDP-ABS.xlsx", sheet = "Data1")[-c(1:9), c(1,63)]
names(gdp) = c("Date", "GDP")
gdp = gdp %>% mutate_if(is.character, as.numeric)
gdp$Date = as.Date(gdp$Date, origin = "1899-12-30")

# match the time window 
gdp = gdp %>% filter(Date <= end_date & Date >= start_index_date)

# match the frequency 
gdp = gdp %>% mutate(Qtr = as.yearqtr(Date)) %>% dplyr::select(-Date)
all_data_original$GDP = gdp$GDP[-1]

# match the unit 
gdp = gdp %>% mutate(gdp_rates = log(GDP) - log(dplyr::lag(GDP, n = 1))) 
gdp = gdp[-1,]

# declare as a ts object 
gdp = ts(gdp$gdp_rates)

#####################################
# CPI: Australia; seasonally adjusted
# import 
cpi = read_excel("data/CPI-ABS.xlsx", sheet = "Data1")[-c(1:9), c(1,10)]
names(cpi) = c("Date", "CPI")
cpi = cpi %>% mutate_if(is.character, as.numeric)
cpi$Date = as.Date(cpi$Date, origin = "1899-12-30")

# match the time window
cpi = cpi %>% filter(Date <= end_date & Date >= start_index_date)

# match the frequency
cpi = cpi %>% mutate(Qtr = as.yearqtr(Date)) %>% dplyr::select(-Date)
all_data_original$CPI = cpi$CPI[-1]

# match the unit
cpi = cpi %>% mutate(cpi_rate = log(CPI) - log(dplyr::lag(CPI, n = 1)))
cpi = cpi[-1,]

# declare as a ts object
cpi = ts(cpi$cpi_rate)

##################################################################
# Non-financial corporate BBB bonds – yield – 10 year target tenor
# import 
# corpBBB_10y_spread = read_excel("data/corporate_bonds-RBA.xls", sheet = "Data")[-c(1:10), c(1,36)]
# names(corpBBB_10y_spread) = c("Date", "yield")
# corpBBB_10y_spread = corpBBB_10y_spread %>% mutate_if(is.character, as.numeric)
# corpBBB_10y_spread$Date = as.Date(corpBBB_10y_spread$Date, origin = "1899-12-30")
# 
# # match the time window
# corpBBB_10y_spread = corpBBB_10y_spread %>% filter(Date <= end_date & Date >= start_pc_date)
# 
# # match the frequency
# corpBBB_10y_spread = corpBBB_10y_spread %>% mutate(Qtr = as.yearqtr(Date)) %>% group_by(Qtr) %>% summarise(yield_qtr = mean(yield))
# 
# # match the unit
# corpBBB_10y_spread$spread_qtr = corpBBB_10y_spread$yield_qtr - zcp3m
# 
# # declare as a ts object
# corpBBB_10y_spread = ts(corpBBB_10y_spread$spread_qtr)

#################################
# S&P/ASX200: daily closing price
# import 
# asx200_new = read_excel("data/ASX200_from2001.xlsx")[, 1:2]
# asx200_new = clean_names(asx200_new)
# 
# asx200_oldest = read_excel("data/ASX200_till2001Q2.xlsx")[,c(1,6)]
# names(asx200_oldest) = c("exchange_date", "close")
# asx200_oldest = asx200_oldest %>% mutate_if(is.character, as.numeric)
# asx200_oldest$exchange_date = as.Date(asx200_oldest$exchange_date, origin = "1899-12-30")
# asx200_oldest = asx200_oldest[order(asx200_oldest$exchange_date),]
# 
# asx200_old = read_excel("data/ASX200_till2001Nov19.xlsx")[,c(1,6)]
# names(asx200_old) = c("exchange_date", "close")
# asx200_old = asx200_old %>% mutate_if(is.character, as.numeric)
# asx200_old$exchange_date = as.Date(asx200_old$exchange_date, origin = "1899-12-30")
# asx200_old = asx200_old[order(asx200_old$exchange_date),]

# asx200 = rbind(asx200_oldest, asx200_old, asx200_new)

asx200 = read.csv("data/asx200_price.csv")[, 1:2]
names(asx200) = c("exchange_date", "close")

# match the time window 
asx200 = asx200 %>% filter(exchange_date <= end_date & exchange_date >= start_index_date)
asx200$exchange_date = as.Date(asx200$exchange_date)

# match the frequency 
asx200 = asx200 %>%
    mutate(Year = year(exchange_date), Quarter = quarter(exchange_date)) %>%
    group_by(Year, Quarter) %>%
    filter(exchange_date == max(exchange_date)) %>%
    ungroup() %>%
    mutate(Qtr = as.yearqtr(exchange_date)) %>%
    dplyr::select(Qtr, close) 
all_data_original$ASX200 = asx200$close[-1]

# match the unit 
asx200 = asx200 %>% mutate(close_qtr_rate = log(close) - log(dplyr::lag(close, n = 1))) 
asx200 = asx200[-1,]

# declare as a ts object 
asx200 = ts(asx200$close_qtr_rate)

############################################
# fixed margin
# unemployment rate NSW; seasonally adjusted
# import
# unemploy_nsw = read_excel("data/Unemployment_rate.xls", sheet = "Data1")[-c(1:9), c(1, 48)]

unemploy_nsw = read.csv("data/Unemployment_rate.csv")
names(unemploy_nsw) = c("Date", "Unemployment_rate")
# unemploy_nsw = unemploy_nsw %>% mutate_if(is.character, as.numeric)
# unemploy_nsw$Date = as.Date(unemploy_nsw$Date, origin = "1899-12-30")
# unemploy_nsw$Date = as.Date(paste0("01-", unemploy_nsw$Date), format = "%d-%b-%Y")

unemploy_nsw$Date = as.Date(unemploy_nsw$Date, format = "%d/%m/%Y")

# match the time window
unemploy_nsw = unemploy_nsw %>% filter(Date <= end_date & Date >= start_pc_date)

# match the frequency
unemploy_nsw = unemploy_nsw %>% mutate(Qtr = as.yearqtr(Date)) %>% group_by(Qtr) %>% summarise(unemploy_qtr_rate = mean(Unemployment_rate))
all_data_original$unemploy = unemploy_nsw$unemploy_qtr_rate

# match the unit

# declare as a ts object
unemploy_nsw = ts(unemploy_nsw$unemploy_qtr_rate)

###############################################
# forex: Australian Dollar Trade-weighted Index
# import 
# forex1 = read_excel("data/Forex_till2010.xls")[-c(1:20), c(1,15)]
# forex2 = read_excel("data/Forex_from2010.xls")[-c(1:10), c(1,3)]
# names(forex1) = c("Date", "AUD_index")
# names(forex2) = c("Date", "AUD_index")
# forex = rbind(forex1, forex2)
# forex = forex %>% mutate_if(is.character, as.numeric)
# forex$Date = as.Date(forex$Date, origin = "1899-12-30")
# rm(forex1, forex2)

forex = read.csv("data/AUD_index.csv")
forex$Date = as.Date(forex$Date, format = "%d/%m/%Y")

# match the time window 
forex = forex %>% filter(Date <= end_date & Date >= start_index_date)

# match the frequency
forex = forex %>%
    mutate(Year = year(Date), Quarter = quarter(Date)) %>%
    group_by(Year, Quarter) %>%
    filter(Date == max(Date)) %>%
    ungroup() %>%
    mutate(Qtr = as.yearqtr(Date)) %>%
    dplyr::select(Qtr, AUD_index)
#forex = forex %>% mutate(Qtr = as.yearqtr(Date)) %>% group_by(Qtr) %>% summarise(AUD_qtr_index = mean(AUD_index)) 
# all_data_original$forex = forex$AUD_qtr_index
all_data_original$forex = forex$AUD_index[-1]

# match the unit 
forex = forex %>% mutate(AUD_qtr_rate = log(AUD_index) - log(dplyr::lag(AUD_index, n = 1))) 
forex = forex[-1,]

# declare as a ts object 
forex = ts(forex$AUD_qtr_rate)
```

## Analyse the time series characteristics

### Raw data plots

```{r}
##################
# individual plots 
par(mfrow = c(3,3))
# 3-month zero-coupon yields
plot(all_data_original$Date, all_data_original$zcp3m, "l", ylab = "(%)", xlab = "Year", main = TeX("r^{(1)}"))
# 10-year zero-coupon yield
plot(all_data_original$Date, all_data_original$zcp10y, "l", ylab = "(%)", xlab = "Year", main = TeX("r^{(40)}"))
# 10-year zero-coupon spread 
plot(all_data_original$Date, zcp10y_spread, "l", ylab = "(%)", xlab = "Year", main = TeX("r^{(40)} - r^{(1)}"))
abline(h = 0, lty = 2)
# Hedonic home value index \n Sydney houses
plot(all_data_original$Date, all_data_original$home_index, "l", ylab = "Index Value", xlab = "Year", main = TeX("HVI"))
# GDP
plot(all_data_original$Date, all_data_original$GDP, "l", ylab = "(%)", xlab = "Year", main = TeX("GDP "))
# CPI
plot(all_data_original$Date, all_data_original$CPI, "l", ylab = "(%)", xlab = "Year", main = TeX("CPI "))
# S&P/ASX200 closing price
plot(all_data_original$Date, all_data_original$ASX200, "l", ylab = "Price", xlab = "Year", main = TeX("ASX"))
# NSW unemployment rates
# plot(all_data_original$Date, all_data_original$unemploy, "l", ylab = "(%)", xlab = "Year", main = "UE")
# AUD trade-weighted index
plot(all_data_original$Date, all_data_original$forex, "l", ylab = "Index Value", xlab = "Year", main = TeX("AUD"))
```

### Transformed data plots

```{r, message=FALSE, warning=FALSE}
##################
# individual plots 
all_data = cbind(zcp3m, zcp10y_spread, home_value_houses_nsw, gdp, asx200, forex)
par(mfrow = c(2,2))
# plot(zcp10y$Qtr, zcp10y_spread, "l", ylab = "spread", xlab = "", main = "10-year zero-coupon spread")
plot(zcp10y$Qtr, home_value_houses_nsw, "l", ylab = "(%)", xlab = "Year", main = TeX("h_t"))
abline(h = 0, lty = 2)
# plot(zcp10y$Qtr, rental_houses_nsw, "l", ylab = "yield", xlab = "", main = "Hedonic rental yields \n Sydney houses")
plot(zcp10y$Qtr, gdp, "l", ylab = "(%)", xlab = "Year", main = TeX("g_t "))
abline(h = 0, lty = 2)
# plot(zcp10y$Qtr, corpBBB_10y_spread, "l")
plot(zcp10y$Qtr, asx200, "l", ylab = "(%)", xlab = "Year", main = TeX("a_t"))
abline(h = 0, lty = 2)
# plot(zcp10y$Qtr, unemploy_nsw, "l", ylab = "rate", xlab = "", main = "NSW unemployment rates")
plot(zcp10y$Qtr, forex, "l", ylab = "(%)", xlab = "Year", main = TeX("d_t"))
abline(h = 0, lty = 2)

##########
# combined
par(mfrow = c(1,2))
# zcp3m vs mortgage
plot(all_data_original$Date, all_data_original$zcp3m, "l", ylab = "(%)", ylim = c(-1,11), xlab = "Year", main = "Correlation of 3-month zero-coupon yield \n and mortgage rate")
lines(all_data_original$Date, lending, "l", lty = 2)
legend(1992.5, 2.5, legend = c(TeX("r^{(1)}"), "MR"), lty = 1:2, box.lty = 0)

# GDP vs CPI
# plot(all_data_original$Date, gdp, "l", ylab = "(%)", ylim = c(-0.08,0.04), xlab = "Year", main = "")
# lines(all_data_original$Date, cpi, "l", lty = 2)
# legend(1992.5, -0.065, legend = c(TeX("g_t"), TeX("c_t")), lty = 1:2, box.lty = 0)

# unemployment vs zcp10y
plot(all_data_original$Date, zcp10y_spread, "l", ylab = "(%)", ylim = c(-1.5,11), xlab = "Year", main = "Correlation of 10-year zero-coupon spread \n and NSW unemployment rate")
lines(all_data_original$Date, unemploy_nsw, "l", lty = 2)
legend(2006, 11, legend = c(TeX("r^{(40)} - r^{(1)}"), TeX("UE")), lty = 1:2, box.lty = 0)
```

### Correlation

- `Mortgage rate` vs `3-month zero-coupon yield` ($\rho$ = 0.90). Mortgage rate removed.

- `10-year zero-coupon spread` vs `NSW unemployment rate` ($\rho$ = 0.75). Unemployment rate removed.

- `10-year zero-coupon spread` vs `10-year BBB-rated corporate bond spread` ($\rho$ = 0.621). Corporate bond was removed.

- `S&P/ASX200 growth` vs `AUD trade index growth` ($\rho$ = 0.42). Both kept.

```{r, message=FALSE, warning=FALSE,fig.width = 5, fig.height = 4, fig.align = "center"}
all_data = cbind(zcp3m, zcp10y_spread, lending, home_value_houses_nsw, gdp, cpi, asx200, unemploy_nsw, forex)
######
# corr 
par(mfrow = c(1,1))
corrplot(cor(all_data, method = "pearson"), tl.col = "black", diag = F, type = "upper", tl.srt = 90)
```

### Principal Component Analysis

On the 1st and 2nd principal components:

- `GDP`, `S&P/ASX200`, `AUD trade index growth` are similar.

- `NSW unemployment rate`, `Sydney home value index growth`, `10-year zero-coupon spread` are similar.

- 6 out of 8 variables explain 90% of variation.

```{r, message=FALSE, warning=FALSE, fig.align = "center"}
set.seed(1)
all_pca = prcomp(all_data, scale = T)

par(mfrow = c(1,2))
biplot(all_pca)

# proportion of variance explained 
pve = (all_pca$sdev ^ 2 / sum(all_pca$sdev ^ 2))
plot(cumsum(pve), xlab = "#Principal Components", main = "Proportion of Variance Explained", ylab = "Cumulative %", ylim = c(0,1), type = "b")
all_data = cbind(zcp3m, zcp10y_spread, home_value_houses_nsw,  gdp, cpi, asx200, forex)
```

# Stationarity for individual variables

## Statistical tests for stationarity

- Augmented Dickey–Fuller test ($H_0$: unit root present in the series).

- Phillips-Perron test ($H_0$: integrated of order 1).

- Significant level is chosen to be 0.05. pvalue = 0.01 means \<= 0.01. No laps at this stage.

- Series to be transformed are: `3-month zero-coupon yield`, `Hedonic rental yields - Sydney houses`.

```{r, message=FALSE, warning=FALSE, results='hide'}
###############
# ACF/PACF plots
par(mfrow = c(2,4))
for (i in 1:length(all_var_names)) {
  acf(all_data[,i])
  pacf(all_data[,i])
}

sig_level = 0.05 # pvalue threshold 


########################
# choose the correct lag 
# PP tests
adf_test_lag = function (x, sig_level = 0.05) {
    output = data.frame(stats = NA, pvalue = NA)
    results = adf.test(x)$type1
    i = 1
    output$stats = results[i,2]
    output$pvalue = results[i,3]
    while (output$pvalue > sig_level & i <= 4) {
        i = i + 1
        output$stats = results[i,2]
        output$pvalue = results[i,3]
    }
    return (output)
}



##############################
# Augmented Dickey–Fuller test
adf_results = data.frame(ADF_statistics = rep(NA,length(all_var_names)), ADF_pvalues = rep(NA,length(all_var_names)), row.names = all_var_names)
for (i in 1:length(all_var_names)) {
  results = adf_test_lag(all_data[,i])
  adf_results[i,1] = results$stats 
  adf_results[i,2] = results$pvalue
} 
adf_results = adf_results %>% mutate(ADF_stationary = ifelse(ADF_pvalues <= sig_level, "Yes", "No evidence"))

######################
# Phillips-Perron test 
pp_results = data.frame(PP_statistics = rep(NA, length(all_var_names)), PP_pvalues = rep(NA, length(all_var_names)), row.names = all_var_names)
for (i in 1:ncol(all_data)) {
  results = aTSA::pp.test(all_data[,i])
  pp_results[i,1] = results[1,2]
  pp_results[i,2] = results[1,3]
} 
pp_results = pp_results %>% mutate(PP_stationary = ifelse(PP_pvalues <= sig_level, "Yes", "No evidence"))
```

```{r, include=TRUE}
########################
# vars to be transformed
non_stationary_var_names = c("3-month zero-coupon yield")
pp_results_trans = data.frame(PP_statistics = rep(NA, length(non_stationary_var_names)), PP_pvalues = rep(NA, length(non_stationary_var_names)), row.names = non_stationary_var_names)
adf_results_trans = data.frame(ADF_statistics = rep(NA, length(non_stationary_var_names)), ADF_pvalues = rep(NA, length(non_stationary_var_names)), row.names = non_stationary_var_names)

###########################
# printing the prev results 
# pvalue = 0.01 actually means pvalue < 0.01
print_pvalue = function (x) {
    # input is a vector of pvalues 
    output = ifelse(x == 0.01, "< 0.01", x)
}

# print results 
stationary_results_lag4 = cbind(adf_results, pp_results)
stationary_results_lag4$ADF_pvalues = print_pvalue(stationary_results_lag4$ADF_pvalues)
stationary_results_lag4$PP_pvalues = print_pvalue(stationary_results_lag4$PP_pvalues)
kbl(stationary_results_lag4) %>% kable_styling(latex_options = c("striped", "scale_down"))
```

## Transform if non-stationary

Non-stationary series were differenced once. Now, all series starts from Q2 2002.

```{r, results='hide'}
###########################
# 3-month zero-coupon yield
# differencing 
#zcp3m_stationary = zcp3m - dplyr::lag(as.vector(zcp3m), n = 1)
zcp3m_stationary = zcp3m
zcp3m_stationary = ts(zcp3m_stationary[-1])

# store new stationarity test results 
pp_results_trans[1,1] = tseries::pp.test(zcp3m_stationary)$statistic
pp_results_trans[1,2] = tseries::pp.test(zcp3m_stationary)$p.value
results = adf.test(zcp3m_stationary)$type1
adf_results_trans[1,1] = results[1,2] # stats
adf_results_trans[1,2] = results[1,3] # pvalue 

############################
# 10-year zero-coupon spread
# differencing 
# zcp10y_spread_stationary = zcp10y_spread - dplyr::lag(as.vector(zcp10y_spread), n = 1)
# zcp10y_spread_stationary = ts(zcp10y_spread_stationary[-1])
# 
# # store new stationarity test results 
# pp_results_trans[2,1] = tseries::pp.test(zcp10y_spread_stationary)$statistic
# pp_results_trans[2,2] = tseries::pp.test(zcp10y_spread_stationary)$p.value
# results = adf.test(zcp10y_spread_stationary)$type1
# adf_results_trans[2,1] = results[1,2] # stats
# adf_results_trans[2,2] = results[1,3] # pvalue 

# #######################################
# # Hedonic rental yields - NSW houses
# # differencing 
# rental_houses_nsw_stationary = rental_houses_nsw - dplyr::lag(as.vector(rental_houses_nsw), n = 1)
# rental_houses_nsw_stationary = ts(rental_houses_nsw_stationary[-1])

# # store new stationarity test results 
# results = tseries::pp.test(rental_houses_nsw_stationary)
# pp_results_trans[2,1] = results$statistic
# pp_results_trans[2,2] = results$p.value
# results = adf.test(rental_houses_nsw_stationary)$type1
# adf_results_trans[2,1] = results[1,2] # stats
# adf_results_trans[2,2] = results[1,3] # pvalue 


#########################
# Unemployment rate - NSW
# # differencing 
# unemploy_nsw_stationary = unemploy_nsw - dplyr::lag(as.vector(unemploy_nsw), n = 1)
# unemploy_nsw_stationary = ts(unemploy_nsw_stationary[-1])
# 
# # store new stationarity test results 
# results = tseries::pp.test(unemploy_nsw_stationary)
# pp_results_trans[4,1] = results$statistic
# pp_results_trans[4,2] = results$p.value
# results = adf.test(unemploy_nsw_stationary)$type1
# adf_results_trans[4,1] = results[1,2] # stats
# adf_results_trans[4,2] = results[1,3] # pvalue 

####################
# match time windows
zcp10y_spread = ts(zcp10y_spread[-1])
home_value_houses_nsw = ts(home_value_houses_nsw[-1])
gdp = ts(gdp[-1])
cpi = ts(cpi[-1])
asx200 = ts(asx200[-1])
forex = ts(forex[-1])
```

```{r}
pp_results_trans = pp_results_trans %>% mutate(PP_stationary = ifelse(PP_pvalues <= 0.05, "Yes", "No evidence"))
adf_results_trans = adf_results_trans %>% mutate(ADF_stationary = ifelse(ADF_pvalues <= 0.05, "Yes", "No evidence"))

stationary_results_lag1 = cbind(adf_results_trans, pp_results_trans)
stationary_results_lag1$ADF_pvalues = print_pvalue(stationary_results_lag1$ADF_pvalues)
stationary_results_lag1$PP_pvalues = print_pvalue(stationary_results_lag1$PP_pvalues)
kbl(stationary_results_lag1) %>% kable_styling(latex_options = c("striped", "scale_down"))
```

# Full VAR

## Find the optimal order p

Notes on R built-in functions:

- Formulas for the statistics ([link](http://www.phdeconomics.sssup.it/documents/Lesson18.pdf)). Ouput `HQ` is HQC, `SIC` is BIC, `FPE` is final prediction error.

- `VARSelect` for every lag up to maximum are based on the same sample size, which is not suitable. We compute the statistics mannually.

- `AIC()`, `BIC()` do not use the right residual covariance.

Three types of VAR deterministic regressors:

- `none`: optimal lag is 1.

  \begin{align*}
    \mathbf{y}_{t} = \boldsymbol{\beta}_{1}\mathbf{y}_{t-1}+ \boldsymbol{\beta}_{2}\mathbf{y}_{t-2}+\cdots++ \boldsymbol{\beta}_{p}\mathbf{y}_{t-p} + \boldsymbol{\epsilon}. \end{align*}

- `const`: optimal lag is 1.

\begin{align*}
    \mathbf{y}_{t} = \boldsymbol{\alpha} + \boldsymbol{\beta}_{1}\mathbf{y}_{t-1}+ \boldsymbol{\beta}_{2}\mathbf{y}_{t-2}+\cdots++ \boldsymbol{\beta}_{p}\mathbf{y}_{t-p} + \boldsymbol{\epsilon}. \end{align*}

- `trend`: optimal lag is 1.

\begin{align*}
    \mathbf{y}_{t} = \boldsymbol{\alpha} + \vect \gamma t + \boldsymbol{\beta}_{1}\mathbf{y}_{t-1}+ \boldsymbol{\beta}_{2}\mathbf{y}_{t-2}+\cdots++ \boldsymbol{\beta}_{p}\mathbf{y}_{t-p} + \boldsymbol{\epsilon}. \end{align*}

- `both`: `const` and `trend`. optimal lag is 1.

```{r}
all_data_stationary = cbind(zcp3m_stationary, zcp10y_spread, home_value_houses_nsw, gdp, cpi, asx200, forex)
all_data_stationary[,1] = all_data_stationary[,1]/100
all_data_stationary[,2] = all_data_stationary[,2]/100
# 1. Determine the optimal lag length (testing up to 8 periods)
# type = "const" is standard for stationary data with non-zero means
lag_select <- VARselect(all_data_stationary, lag.max = 8, type = "const")

# View the recommendations from different information criteria
print(lag_select$selection)

# 2. Extract the optimal lag order according to the Akaike Information Criterion (AIC)
# Alternatively, use "SC(n)" for the stricter Schwarz/BIC criterion
p_opt <- lag_select$selection["AIC(n)"]

# 3. Fit the VAR model
var_model <- VAR(all_data_stationary, p = p_opt, type = "const")

# 4. Display the estimation results
summary(var_model)

# Forecast the next 12 periods (e.g., 12 quarters = 3 years)
# ci = 0.90 sets a 90% confidence interval for the forecast bands
var_forecast <- predict(var_model, n.ahead = 60, ci = 0.90)
plot(var_forecast[["fcst"]][["zcp3m_stationary"]][,1])


# Simulation
# 1. Extract VAR(1) coefficients and constant
A1 <- Acoef(var_model)[[1]]
c_vec <- Bcoef(var_model)[, "const"]

# 2. Extract the residual covariance matrix (Sigma)
Sigma <- summary(var_model)$covres

# 3. Get the last observed row of your stationary data to start the simulation
last_obs <- as.numeric(tail(all_data_stationary, 1))
n_vars <- length(last_obs)

# Simulation parameters
n_sims <- 100      # Number of stochastic paths to generate
horizon <- 10       # Number of quarters to forecast

# Initialize a 3D array to store the results
# Dimensions: [Time (horizon), Variables, Simulation Path]
sim_paths <- array(NA, dim = c(horizon, n_vars, n_sims))

# Run the simulations
set.seed(123) # For reproducibility
for (i in 1:n_sims) {
  
  # Reset the current state to the last known historical observation
  current_state <- last_obs
  
  for (t in 1:horizon) {
    # 1. Draw a random shock vector from the multivariate normal distribution
    shock <- mvrnorm(n = 1, mu = rep(0, n_vars), Sigma = Sigma)
    
    # 2. Calculate the next period's values
    next_state <- c_vec + as.vector(A1 %*% current_state) + shock
    
    # 3. (Optional) Apply bounds here if you want to restrict extreme shocks
    # e.g., capping the interest rate difference if necessary
    
    # 4. Store the result and roll the state forward
    sim_paths[t, , i] <- next_state
    current_state <- next_state
  }
}

# 1. Set up the plotting grid
# For 7 variables, a 3x3 grid (with two empty panels) is optimal
par(mfrow = c(3, 3), 
    mar = c(4, 4, 3, 1),  # Inner margins for each plot (bottom, left, top, right)
    oma = c(2, 0, 2, 0))  # Outer margins for the whole figure

# 2. Define mathematical titles for the panels (matching your paper's notation)
plot_titles <- list(
  zcp3m_stationary = expression(r^{(1)}),
  zcp10y_spread = expression(r^{(40)} - r^{(1)}),
  home_value_houses_nsw = expression(h[t]),
  gdp = expression(g[t]),
  cpi = expression(pi[t]),
  asx200 = expression(s[t]),
  forex = expression(e[t])
)

# 3. Extract the variable names from your data
var_names <- colnames(all_data_stationary)
n_hist <- nrow(all_data_stationary)

# 4. Loop through each variable to create its panel
for (v in var_names) {
  
  # Extract historical data
  hist_y <- all_data_stationary[, v]
  
  # Extract forecast data
  fc_mean <- var_forecast$fcst[[v]][, "fcst"]
  fc_lower <- var_forecast$fcst[[v]][, "lower"]
  fc_upper <- var_forecast$fcst[[v]][, "upper"]
  
  # To avoid a gap on the chart, stitch the last historical point to the start of the forecast
  last_hist_val <- tail(hist_y, 1)
  fc_mean_line <- c(last_hist_val, fc_mean)
  fc_lower_line <- c(last_hist_val, fc_lower)
  fc_upper_line <- c(last_hist_val, fc_upper)
  
  # Create time indices
  n_fcst <- length(fc_mean)
  t_hist <- 1:n_hist
  t_fcst <- n_hist:(n_hist + n_fcst) # Starts at the last historical period
  
  # Determine optimal Y-axis limits to fit everything
  ylim_max <- max(c(hist_y, fc_upper_line), na.rm = TRUE)
  ylim_min <- min(c(hist_y, fc_lower_line), na.rm = TRUE)
  
  # Plot the historical data
  plot(t_hist, hist_y, 
       type = "l", 
       lty = 1,              # Solid line
       col = "black", 
       ylim = c(ylim_min, ylim_max),
       xlim = c(1, n_hist + n_fcst),
       xlab = "Quarters", 
       main = plot_titles[[v]], 
       cex.main = 1.2)
  
  # Add the mean forecast (solid line)
  lines(t_fcst, fc_mean_line, lty = 1, col = "black")
  
  # Add the 90% Confidence Intervals (dotted lines)
  lines(t_fcst, fc_lower_line, lty = 3, col = "black")
  lines(t_fcst, fc_upper_line, lty = 3, col = "black")
}
```
